---
title: "Machine Learning Project - Building a model to predict exercise manner"
author: "S.N"
date: "Sunday, March 22, 2015"
output: html_document
---

##Purpose
To predict the manner in which 6 participants have exercised based on the given dataset generated from accelerometers on the belt, forearm, arm, and dumbell of the participants. There are five different possible manners, 1 is the correct manner while the other 4 are incorrect: A,B,C,D,E. This is 'classe' variable in the dataset which will be the outcome while the other variables in the dataset are predictors

##Method
1. Tidy the data set

A number of variables have empty observations and DIV/0. This is replaced by entering NA in lieu of these observations. There are some variables which have no observations at all. These will be removed for the training dataset.

```{r eval=FALSE}
#Load data
training <- read.csv(file="pml-training.csv",header=TRUE)
testing <- read.csv(file="pml-testing.csv",header=TRUE)

#clean data remvove NA and DIV/0
training[training == ""] <- NA
training[training == "#DIV/0!"] <- NA
training <- training[,colSums(is.na(training)) == 0]
```

2. The dataset is tidied further by removing the variables "X","user_name" and timestamps. Variables which are zero covariates are also removed.

```{r eval=FALSE}
#Remove zero covariates
nearZeroVar(training,saveMetrics=TRUE)
#new_window variable can be removed from above

#remove X, username and timestamps
training_subset <- training[,c(-1,-2,-3,-4,-5,-6)]
````

3. Partition data into mini dataset for exploratory purpose.

```{r eval=FALSE}
#split training set into smaller set 
inTrain <- createDataPartition(y=training_subset$classe,p=0.05,list=FALSE)
trainingvalidate <- training_subset[-inTrain,]
trainingsplit <- training_subset[inTrain,]
```

4. Create exploratory model
```{r eval=FALSE}
fitmodel <- train(classe ~ .,data=trainingsplit,method="rf",prox=TRUE,trControl=trainControl(preProcOptions=list(k=3)))

```

5. Find the top important predictors.
```{r eval=FALSE}
#find important variables
varImp(fitmodel)
```

6. Create bigger training set. Because of computing restraints the raw data is partitioned to 40% for training set.

```{r eval=FALSE}
#split raw data into training set and validation set  
inTrain <- createDataPartition(y=training_subset$classe,p=0.40,list=FALSE)
trainingvalidate <- training_subset[-inTrain,]
trainingsplit <- training_subset[inTrain,]
```

6. Fine tune the model by repeating step 4 and 5 and removing unimportant predictor. The final chosen predictor as below:
```{r eval=FALSE}
#Chosen model
fitmodel <- train(classe ~ num_window + roll_belt + pitch_forearm + magnet_dumbbell_z + magnet_dumbbell_y + yaw_belt + pitch_belt,data=trainingsplit,method="rf",prox=TRUE,trControl=trainControl(preProcOptions=list(k=3)))
```

7. Run on validation set
```{r eval=FALSE}
#Predict validation dataset with training model
predvalidate <- predict(fitmodel,trainingvalidate)
#See how many mispredicted
table(predvalidate,trainingvalidate$classe)
```
+--------------+-----+-----+-----+-----+----+
|predvalidate  |  A  |  B  |  C  |  D  |  E |
+--------------+-----+-----+-----+-----+----+
|A             |3345 |  1  | 0   |  0  |  0 | 
+--------------+-----+-----+-----+-----+----+
|B             |3    |2264 |4    |2    |4   |
+--------------+-----+-----+-----+-----+----+
|C             |0    |12   |2049 |  3  |  0 |
+--------------+-----+-----+-----+-----+----+
|D             |0    |1    |0    |1924 |  5 |
+--------------+-----+-----+-----+-----+----+
|E             |0    |  0  |  0  |   0 |2155|
+--------------+-----+-----+-----+-----+----+

From the table above 0.17 percent of the outcome has been mispredicted.

The accuracy of the chosen model is 99 percent:

**Random Forest**

**7850 samples**
  **53 predictor**
   **5 classes: 'A', 'B', 'C', 'D', 'E'** 

**No pre-processing**
**Resampling: Bootstrapped (25 reps)** 

**Summary of sample sizes: 7850, 7850, 7850, 7850, 7850, 7850, ...** 

**Resampling results across tuning parameters:**

+-----------+---------------+---------------+------------------+-----------------+
|  **mtry** |  **Accuracy** |  **Kappa**    |  **Accuracy SD** |  **Kappa SD**   | 
+-----------+---------------+---------------+------------------+-----------------+
|  **2**    | **0.9932145** |**0.9914094**  |  **0.001642673** | **0.002080360** |
+-----------+---------------+---------------+------------------+-----------------+
|  **4**    | **0.9926201** | **0.9906575** | **0.002101274**  | **0.002657833** |
+-----------+---------------+---------------+------------------+-----------------+
|  **7**    | **0.9885686** | **0.9855256** | **0.003106409**  | **0.003936698** |
+-----------+---------------+---------------+------------------+-----------------+

**Accuracy was used to select the optimal model using  the largest value.**
**The final value used for the model was mtry = 2. ** 

8. Run on testing set to predict exercise manner
```{r eval=FALSE}
answers <- predict(fitmodel,testing)
```

9. Submit answer to coursera website.